Supplementary materials: $R^*$ convergence diagnostic

### 2020-11-09 src/s_trends.Rmd
cleaning up repo and rerunning rhat experiments

---
title: "Examining the effect of a trend on R*"
output: html_notebook
---

@@

```{r}
res <- readRDS("../output/experiment_trends_all_dim.rds")
g <-
  res %>%
  select(iters, trend, Rstar, Rstar_split, method) %>%
  melt(id.vars=c("iters", "trend", "method")) %>%
  mutate(variable=case_when(variable=="Rstar"~"A. R*",
                            variable=="Rstar_split"~"B. Split-R*")) %>%
  mutate(method=as.character(method)) %>% 
  mutate(method=if_else(method=="gbm", "GBM", "RF")) %>% 
  ggplot(aes(y = value, x = trend)) +
  geom_jitter(aes(colour=as.factor(method), shape=as.factor(method))) +
  facet_grid(vars(iters), vars(variable)) +
  geom_hline(yintercept = 1) +
  theme(strip.text.x = element_text(size=16, colour="black"),
        strip.text.y = element_text(size=16, colour="black"),
        axis.text.x = element_text(size=16, colour="black"),
        axis.text.y = element_text(size=16, colour="black"),
        axis.title = element_text(size=16, colour="black"),
        legend.text = element_text(size=16, colour="black"),
        legend.title = element_text(size=16, colour="black")) +
  xlab("Trend") +
  ylab("Value") +
  scale_colour_grey("Classifier") +
  scale_shape("Classifier")
ggsave("../output/trends_all_dim.pdf", g, width = 8, height = 6)
```

@@

```{r}
g <- readRDS("../output/experiment_trends_one_dim.rds") %>%
  select(trend, iters, rep, dimension, Rstar_split, method) %>%
  unique() %>%
  select(-rep) %>%
  melt(c("trend", "iters", "dimension", "method")) %>%
  filter(variable=="Rstar_split") %>%
  mutate(method=as.character(method)) %>% 
  mutate(method=if_else(method=="gbm", "GBM", "RF")) %>% 
  ggplot(aes(y = value, x = trend)) +
  geom_jitter(aes(colour=as.factor(method),
                  shape=as.factor(method))) +
  facet_grid(vars(iters), vars(dimension)) +
  geom_hline(yintercept = 1) +
  theme(strip.text.x = element_text(size=16, colour="black"),
        strip.text.y = element_text(size=16, colour="black"),
        axis.text.x = element_text(size=16, colour="black"),
        axis.text.y = element_text(size=16, colour="black"),
        axis.title = element_text(size=16, colour="black"),
        legend.text = element_text(size=16, colour="black"),
        legend.title = element_text(size=16, colour="black")) +
  xlab("Trend") +
  ylab("Value") +
  scale_colour_grey("Classifier") +
  scale_shape("Classifier")
g

@@

```{r}
# remove double iterates for rhat from rfs
g <- readRDS("../output/experiment_trends_joint_distribution.rds") %>%
  as.data.frame() %>%
  select(iters, rho, rep, zsRhat, Rstar_split, method) %>%
  unique() %>%
  select(-rep) %>%
  melt(c("rho", "iters", "method")) %>%
  filter(!((method=="rf")&(variable=="zsRhat"))) %>% 
  mutate(value=as.numeric(as.character(value))) %>% 
  mutate(variable=case_when(variable=="zsRhat"~"A. Split-Rhat",
                            variable=="Rstar_split"~"B. Split-R*")) %>% 
  mutate(iters=as.numeric(as.character(iters))) %>% 
  mutate(method=as.character(method)) %>% 
  mutate(method=if_else(method=="rf", "RF", "GBM")) %>% 
  ggplot(aes(y = value, x = rho)) +
  geom_jitter(aes(colour=as.factor(method),
                  shape=as.factor(method))) +
  facet_grid(vars(iters), vars(variable)) +
  geom_hline(yintercept = 1) +
  theme(strip.text.x = element_text(size=16, colour="black"),
        strip.text.y = element_text(size=16, colour="black"),
        axis.text.x = element_text(size=16, colour="black"),
        axis.text.y = element_text(size=16, colour="black"),
        axis.title = element_text(size=16, colour="black"),
        legend.text = element_text(size=16, colour="black"),
        legend.title = element_text(size=16, colour="black")) +
  ylab("Value") +
  xlab(TeX("$\\rho$"))  +
  scale_colour_grey("Classifier") +
  scale_shape("Classifier")
g

@@

```{r}
res <- readRDS("../output/experiment_trends_rho_nearer1.rds") %>%
  mutate(method=as.character(method)) %>% 
  mutate(method=if_else(method=="gbm", "GBM", "RF"))

# ggplot(data = res, aes(x = rho, y = zsRhat)) +
#   geom_jitter() +
#   facet_grid(vars(iters)) +
#   geom_hline(yintercept = 1) +
#   ggtitle('R* without splitting') +
#   ylim(0, 6)

>>>>

---
title: "Examining the effect of sampler trends on R*"
output: html_notebook
---

@@

Plot
```{r}
res <- readRDS("../output/experiment_trends_all_dim.rds")
g <-
  res %>%
  select(iters, trend, Rstar, Rstar_split, method) %>%
  melt(id.vars=c("iters", "trend", "method")) %>%
  mutate(variable=case_when(variable=="Rstar"~"A. R*",
                            variable=="Rstar_split"~"B. Split-R*")) %>%
  mutate(method=as.character(method)) %>% 
  mutate(method=if_else(method=="gbm", "GBM", "RF")) %>% 
  ggplot(aes(y = value, x = trend)) +
  geom_jitter(aes(colour=as.factor(method), shape=as.factor(method))) +
  facet_grid(vars(iters), vars(variable)) +
  geom_hline(yintercept = 1) +
  theme(strip.text.x = element_text(size=16, colour="black"),
        strip.text.y = element_text(size=16, colour="black"),
        axis.text.x = element_text(size=16, colour="black"),
        axis.text.y = element_text(size=16, colour="black"),
        axis.title = element_text(size=16, colour="black"),
        legend.text = element_text(size=16, colour="black"),
        legend.title = element_text(size=16, colour="black")) +
  xlab("Trend") +
  ylab("Value") +
  scale_colour_grey("Classifier") +
  scale_shape("Classifier")
ggsave("../output/trends_all_dim.pdf", g, width = 8, height = 6)
```

@@

Plot
```{r}
g <- readRDS("../output/experiment_trends_one_dim.rds") %>%
  select(trend, iters, rep, dimension, Rstar_split, method) %>%
  unique() %>%
  select(-rep) %>%
  melt(c("trend", "iters", "dimension", "method")) %>%
  filter(variable=="Rstar_split") %>%
  mutate(method=as.character(method)) %>% 
  mutate(method=if_else(method=="gbm", "GBM", "RF")) %>% 
  ggplot(aes(y = value, x = trend)) +
  geom_jitter(aes(colour=as.factor(method),
                  shape=as.factor(method))) +
  facet_grid(vars(iters), vars(dimension)) +
  geom_hline(yintercept = 1) +
  theme(strip.text.x = element_text(size=16, colour="black"),
        strip.text.y = element_text(size=16, colour="black"),
        axis.text.x = element_text(size=16, colour="black"),
        axis.text.y = element_text(size=16, colour="black"),
        axis.title = element_text(size=16, colour="black"),
        legend.text = element_text(size=16, colour="black"),
        legend.title = element_text(size=16, colour="black")) +
  xlab("Trend") +
  ylab("Value") +
  scale_colour_grey("Classifier") +
  scale_shape("Classifier")
g

@@

Plot
```{r}
# remove double iterates for rhat from rfs
g <- readRDS("../output/experiment_trends_joint_distribution.rds") %>%
  as.data.frame() %>%
  dplyr::select(iters, rho, rep, zsRhat, zfsRhat, Rstar_split, method) %>%
  unique() %>%
  dplyr::select(-rep) %>%
  mutate(zsRhat=max(as.numeric(as.character(zsRhat), as.numeric(as.character(zfsRhat))))) %>% 
  dplyr::select(-zfsRhat) %>% 
  melt(c("rho", "iters", "method")) %>%
  filter(!((method=="rf")&(variable=="zsRhat"))) %>% 
  mutate(value=as.numeric(as.character(value))) %>% 
  mutate(variable=case_when(variable=="zsRhat"~"A. Split-Rhat",
                            variable=="Rstar_split"~"B. Split-R*")) %>% 
  mutate(iters=as.numeric(as.character(iters))) %>% 
  mutate(method=as.character(method)) %>% 
  mutate(method=if_else(method=="rf", "RF", "GBM")) %>% 
  ggplot(aes(y = value, x = rho)) +
  geom_jitter(aes(colour=as.factor(method),
                  shape=as.factor(method))) +
  facet_grid(vars(iters), vars(variable)) +
  geom_hline(yintercept = 1) +
  theme(strip.text.x = element_text(size=16, colour="black"),
        strip.text.y = element_text(size=16, colour="black"),
        axis.text.x = element_text(size=16, colour="black"),
        axis.text.y = element_text(size=16, colour="black"),
        axis.title = element_text(size=16, colour="black"),
        legend.text = element_text(size=16, colour="black"),
        legend.title = element_text(size=16, colour="black")) +
  ylab("Value") +
  xlab(TeX("$\\rho$"))  +
  scale_colour_grey("Classifier") +
  scale_shape("Classifier")
g

@@

Plot
```{r}
res <- readRDS("../output/experiment_trends_rho_nearer1.rds") %>%
  mutate(method=as.character(method)) %>% 
  mutate(method=if_else(method=="gbm", "GBM", "RF"))

### 2020-11-09 src/s_wide_distributions.Rmd
cleaning up repo and rerunning rhat experiments

full_data <- rstan::extract(fit, permuted=F, inc_warmup=F)
r_star_gbm <- r_star(full_data, uncertainty = T, method="gbm")
r_star_rf <- r_star(full_data, uncertainty = T, method="rf")

@@

mon <- monitor_extra(split_data(full_data))
mean(mon$zfsRhat>1.01)
mean(mon$zsseff<400)
mean(mon$tailseff<400)

b_df <- tibble(samples_cp=mon$zfsRhat[1:250])

ha <- tibble(bulk_ess=mon$zsseff,
       tail_ess=mon$tailseff,
       rhat=mon$zfsRhat) %>% 
  melt() %>% 
  mutate(variable=case_when(variable=="bulk_ess"~"Bulk ESS",
                            variable=="tail_ess"~"Tail ESS",
                            variable=="rhat"~"Split-Rhat")) %>% 
  ggplot(aes(x=value)) +
  geom_histogram() + 
  facet_wrap(~relevel(as.factor(variable), "Split-Rhat", "Bulk ESS", "Tail ESS"), scales="free") + xlab("Value") +
  ylab("Count") +
  theme(text=element_text(size=12, colour="black"),
        axis.text=element_text(size=12, colour="black"),
        strip.text = element_text(size=12, colour="black"),
        axis.title.x = element_blank())
ha
aa_df <- tibble(bulk_ess=mon$zsseff,
       tail_ess=mon$tailseff,
       rhat=mon$zfsRhat,
       source="Centered")
# ggsave("../output/wide_cp_diagnostics.pdf", ha, width = 12, height = 8)
```

@@

mon <- monitor_extra(split_data(full_data))
mean(mon$zfsRhat>1.01)
mean(mon$zsseff<400)
mean(mon$tailseff<400)

h <- tibble(bulk_ess=mon$zsseff,
       tail_ess=mon$tailseff,
       rhat=mon$zfsRhat) %>% 
  melt() %>% 
  mutate(variable=case_when(variable=="bulk_ess"~"Bulk-ESS",
                            variable=="tail_ess"~"Tail-ESS",
                            variable=="rhat"~"Split-Rhat")) %>% 
  ggplot(aes(x=value)) +
  geom_histogram() + 
  facet_wrap(~relevel(as.factor(variable), "Split-Rhat", "Bulk-ESS", "Tail-ESS"), scales="free") + xlab("Value") +
  ylab("Count") +
  theme(text=element_text(size=12, colour="black"),
        axis.text=element_text(size=12, colour="black"),
        strip.text = element_text(size=12, colour="black"),
        axis.title.x = element_blank())
# ggsave("../output/wide_ncp_diagnostics.pdf", h, width = 12, height = 8)

aa_df <- aa_df %>%
  bind_rows(tibble(bulk_ess=mon$zsseff,
       tail_ess=mon$tailseff,
       rhat=mon$zfsRhat,
       source="Non-centered"))

@@

full_data <- rstan::extract(fit, permuted=F, inc_warmup=F)
r_star_gbm <- r_star(full_data, uncertainty = T, method="gbm")
r_star_rf <- r_star(full_data, uncertainty = T, method="rf")
a_df <- tibble(samples_400_gbm=r_star_gbm,
               samples_400_rf=r_star_rf)
saveRDS(a_df, "../output/mvt_wide_400_summary.rds")
mon <- monitor_extra(split_data(full_data))
mean(mon$zfsRhat>1.01)
mean(mon$zsseff<400)
mean(mon$tailseff<400)
```

@@

a_df <- readRDS("../output/mvt_wide_400_summary.rds")
a_df <- a_df %>% 
    mutate(samples_1000_gbm=r_star_gbm,
           samples_1000_rf=r_star_rf)
saveRDS(a_df, "../output/mvt_wide_summaries.rds")
mon <- monitor_extra(split_data(full_data))
mean(mon$zfsRhat>1.01)
mean(mon$zsseff<400)
mean(mon$tailseff<400)
a_df <- readRDS("../output/mvt_wide_summaries.rds")
a_df_gbm <- a_df %>%
  select(contains("gbm")) %>% 
  rename(samples_400=samples_400_gbm,
         samples_1000=samples_1000_gbm) %>% 
  mutate(Classifier="C. GBM")
a_df_rf <- a_df %>%
  select(contains("rf")) %>% 
  rename(samples_400=samples_400_rf,
         samples_1000=samples_1000_rf) %>% 
  mutate(Classifier="D. RF")

>>>>

full_data <- rstan::extract(fit, permuted=F, inc_warmup=F)

## R* calcs
r_star_gbm <- r_star(full_data, uncertainty = T, method="gbm")
r_star_rf <- r_star(full_data, uncertainty = T, method="rf")

@@

## Rhat cals and ESSs
mon <- monitor_extra(full_data)

f_rhat_max <- function(mon) {
  rhat <- mon$zfsRhat
  rhat1 <- mon$zsRhat
  rhat <- tibble(rhat, rhat1) %>% 
    apply(1, max)
  return(rhat)
}

rhats <- f_rhat_max(mon)
mean(rhats>1.01)
mean(mon$zsseff<400)
mean(mon$tailseff<400)

## plot traditional
ha <- tibble(bulk_ess=mon$zsseff,
       tail_ess=mon$tailseff,
       rhat=rhats) %>% 
  melt() %>% 
  mutate(variable=case_when(variable=="bulk_ess"~"Bulk ESS",
                            variable=="tail_ess"~"Tail ESS",
                            variable=="rhat"~"Split-Rhat")) %>% 
  ggplot(aes(x=value)) +
  geom_histogram() + 
  facet_wrap(~relevel(as.factor(variable), "Split-Rhat", "Bulk ESS", "Tail ESS"), scales="free") + xlab("Value") +
  ylab("Count") +
  theme(text=element_text(size=12, colour="black"),
        axis.text=element_text(size=12, colour="black"),
        strip.text = element_text(size=12, colour="black"),
        axis.title.x = element_blank())
ha
aa_df <- tibble(bulk_ess=mon$zsseff,
       tail_ess=mon$tailseff,
       rhat=rhats,
       source="Centered")
# ggsave("../output/wide_cp_diagnostics.pdf", ha, width = 12, height = 8)
```

@@

mon <- monitor_extra(full_data)
rhats <- f_rhat_max(mon)
mean(rhats>1.01)
mean(mon$zsseff<400)
mean(mon$tailseff<400)

h <- tibble(bulk_ess=mon$zsseff,
       tail_ess=mon$tailseff,
       rhat=rhats) %>% 
  melt() %>% 
  mutate(variable=case_when(variable=="bulk_ess"~"Bulk-ESS",
                            variable=="tail_ess"~"Tail-ESS",
                            variable=="rhat"~"Split-Rhat")) %>% 
  ggplot(aes(x=value)) +
  geom_histogram() + 
  facet_wrap(~relevel(as.factor(variable), "Split-Rhat", "Bulk-ESS", "Tail-ESS"), scales="free") + xlab("Value") +
  ylab("Count") +
  theme(text=element_text(size=12, colour="black"),
        axis.text=element_text(size=12, colour="black"),
        strip.text = element_text(size=12, colour="black"),
        axis.title.x = element_blank())
# ggsave("../output/wide_ncp_diagnostics.pdf", h, width = 12, height = 8)

## plot all traditional measures
aa_df <- aa_df %>%
  bind_rows(tibble(bulk_ess=mon$zsseff,
       tail_ess=mon$tailseff,
       rhat=rhats,
       source="Non-centered"))

@@

full_data <- rstan::extract(fit, permuted=F, inc_warmup=F)
r_star_gbm <- r_star(full_data, uncertainty = T, method="gbm")
r_star_rf <- r_star(full_data, uncertainty = T, method="rf")
a_df <- tibble(samples_400_gbm=r_star_gbm,
               samples_400_rf=r_star_rf)
saveRDS(a_df, "../output/mvt_wide_400_summary.rds")
mon <- monitor_extra(full_data)
rhats <- f_rhat_max(mon)
mean(rhats>1.01)
mean(mon$zsseff<400)
mean(mon$tailseff<400)
```

@@

a_df <- readRDS("../output/mvt_wide_400_summary.rds")
a_df <- a_df %>% 
    mutate(samples_1000_gbm=r_star_gbm,
           samples_1000_rf=r_star_rf)
saveRDS(a_df, "../output/mvt_wide_summaries.rds")
mon <- monitor_extra(full_data)
rhats <- f_rhat_max(mon)
mean(rhats>1.01)
mean(mon$zsseff<400)
mean(mon$tailseff<400)
a_df <- readRDS("../output/mvt_wide_summaries.rds")
a_df_gbm <- a_df %>%
  select(contains("gbm")) %>% 
  rename(samples_400=samples_400_gbm,
         samples_1000=samples_1000_gbm) %>% 
  mutate(Classifier="C. GBM")
a_df_rf <- a_df %>%
  select(contains("rf")) %>% 
  rename(samples_400=samples_400_rf,
         samples_1000=samples_1000_rf) %>% 
  mutate(Classifier="D. RF")

### 2020-11-09 write_up/arxiv_submission_aug_2020/ms_ba_marked.tex
cleaning up repo and rerunning rhat experiments

It is recommended that $\widehat{R}$, like $R^*$, be calculated using split chains. In Fig. \ref{fig:eight_schools}C, we plot $\widehat{R}$ values obtained when using the original 4 chains (horizontal axis) versus those when using the split chains (vertical axis) for the ten parameters in this model; we do this for both the centered and non-centered models. These show that the values of $\widehat{R}$ for the centered model for both the split and unsplit cases were, for all but a single parameter, above 1.01, echoing the results for $R^*$; also, like $R^*$, $\widehat{R}$ is higher for the split case. All parameters for the non-centered models were below 1.01, indicating convergence.

>>>>

It is recommended that $\widehat{R}$, like $R^*$, be calculated using split chains. In Fig. \ref{fig:eight_schools}C, we plot $\widehat{R}$ values obtained when using the original 4 chains (horizontal axis) versus those when using the split chains (vertical axis) for the ten parameters in this model; we do this for both the centered and non-centered models. \textcolor{red}{These show that the values of $\widehat{R}$ for the centered model using the unsplit chains were below 1.01; when using the chains split into two halves, $\widehat{R}>1.01$ for all but a single parameter. All parameters for the non-centered models were below 1.01, indicating convergence.}

### 2020-11-09 write_up/arxiv_submission_aug_2020/supplementary_materials_marked.tex
cleaning up repo and rerunning rhat experiments

\subsection{10,000-dimensional model}
We next consider a more challenging example -- a target distribution with 10,000 dimensions. In this case, we assume independent standard normals for each dimension. \textcolor{red}{In Figs. \ref{fig:mvt_wide_both}C\&D, we plot the $R^*$ distributions from GBM and RF classifiers for two MCMC runs targeting this distribution: one with 400 iterations, the other with 1000.} In all cases, the distributions were right of $R^*=1$, indicating non-convergence. These results were also echoed by rank-normalised split-$\widehat{R}$, with 65\% of dimensions having $\widehat{R}>1.01$ for the 400 iteration case and 19\% for the 1000 iteration case.

@@

For the ovarian model, we show the results in Fig. \ref{fig:ovarian}. In Fig. \ref{fig:ovarian}A, we show the $R^*$ distributions for each model run, which show that, whereas the ``long'' model run has converged, the ``short'' one has yet to do so. Whilst it is harder to discern, this pattern is mirrored in Fig. \ref{fig:ovarian}B, since the long model has $\widehat{R}<1.01$ for all parameters, whereas the short model had 139 parameters where this was not the case. Similarly, in Figs. \ref{fig:ovarian}C and \ref{fig:ovarian}D, which show the bulk-ESS and tail-ESS respectively, it is evident that, for the short model, there remain a few parameters with low effective sample sizes, whereas the long model has more consistent values for this metric.

@@

For each of these three cases, we consider a range of sample sizes across each of the four chains: 1000, 2000, 5000 and 10,000. At each combination of transition probability matrix (for the fourth chain) and sample size, we perform 40 replicates. In each replicate, we calculate both $R^*$ and rank-normalised split-$\widehat{R}$. The results of these experiments are shown in Fig. \ref{fig:discrete}. In the horizontal axes of both panels, we show the sample size. The vertical axis indicates the value of $R^*$ and split-$\widehat{R}$ in panels A and B, respectively. The stacked sub-panels within A and B indicate the transition matrix used for the fourth chain. Since $\boldsymbol{P}_1=\boldsymbol{P}$, the stationary distribution for the other three chains, both diagnostics should tend towards indicating convergence in these cases, and, for both statistics, this was typically the case. For $R^*$, the replicates were centred on $R^*=1$, although the stochasticity of replicates meant that, in some cases, $R^*>1$. Whereas, $\widehat{R}<1$ in all cases. Considering the simulations using $\boldsymbol{P}_2$ for the fourth chain transition probability matrix, $R^*>1$ across a majority of replicates, and this majority increased with sample size. In this case, $\widehat{R}<1$ for all replicates across the various sample sizes. For simulations using $\boldsymbol{P}_3$ for the fourth chain transition probability matrix, $R^*>1$ across all replicates and the magnitude of median $R^*$ increased with sample size; as for the previous case, $\widehat{R}<1$, indicating convergence.

This example hints that $R^*$ may be particularly adept at detecting poor convergence in small discrete parameter spaces.

@@

Collectively, the results of \S\ref{sec:discrete_small} and \S\ref{sec:discrete_large} hint that $R^*$ may be more reliable at detecting poor convergence than $\widehat{R}$, although further empirical investigation is needed.

>>>>

\subsection{10,000-dimensional model}
We next consider a more challenging example -- a target distribution with 10,000 dimensions. In this case, we assume independent standard normals for each dimension. \textcolor{red}{In Figs. \ref{fig:mvt_wide_both}C\&D, we plot the $R^*$ distributions from GBM and RF classifiers for two MCMC runs targeting this distribution: one with 400 iterations, the other with 1000.} In all cases, the distributions were right of $R^*=1$, indicating non-convergence. These results were also echoed by rank-normalised split-$\widehat{R}$, with \textcolor{red}{19\%} of dimensions having $\widehat{R}>1.01$ for the 400 iteration case and \textcolor{red}{3\%} for the 1000 iteration case.

@@

For the ovarian model, we show the results in Fig. \ref{fig:ovarian}. In Fig. \ref{fig:ovarian}A, we show the $R^*$ distributions for each model run, which show that, whereas the ``long'' model run has converged, the ``short'' one has yet to do so. Whilst it is harder to discern, this pattern is mirrored in Fig. \ref{fig:ovarian}B, since the long model has $\widehat{R}<1.01$ for all parameters, whereas the short model had \textcolor{red}{83} parameters where this was not the case. Similarly, in Figs. \ref{fig:ovarian}C and \ref{fig:ovarian}D, which show the bulk-ESS and tail-ESS respectively, it is evident that, for the short model, there remain a few parameters with low effective sample sizes, whereas the long model has more consistent values for this metric.

@@

For each of these three cases, we consider a range of sample sizes across each of the four chains: 1000, 2000, 5000 and 10,000. At each combination of transition probability matrix (for the fourth chain) and sample size, we perform 40 replicates. In each replicate, we calculate both $R^*$ using a GBM classifier and rank-normalised split-$\widehat{R}$. The results of these experiments are shown in Fig. \ref{fig:discrete}. In the horizontal axes of both panels, we show the sample size. The vertical axis indicates the value of $R^*$ and split-$\widehat{R}$ in panels A and B, respectively. The stacked sub-panels within A and B indicate the transition matrix used for the fourth chain. Since $\boldsymbol{P}_1=\boldsymbol{P}$, the stationary distribution for the other three chains, both diagnostics should tend towards indicating convergence in these cases, and, for both statistics, this was typically the case. For $R^*$, the replicates were centred on $R^*=1$, although the stochasticity of replicates meant that, in some cases, $R^*>1$. Whereas, $\widehat{R}$ was similarly close to 1 in all cases. Considering the simulations using $\boldsymbol{P}_2$ for the fourth chain transition probability matrix, $R^*>1$ across a majority of replicates, and this majority increased with sample size. In this case, $1.01>\widehat{R}>1$ for all replicates across the various sample sizes. For simulations using $\boldsymbol{P}_3$ for the fourth chain transition probability matrix, $R^*>1$ across all replicates and the magnitude of median $R^*$ increased with sample size; in this case, $\widehat{R}>1.01$ in all replicates.

This example shows that $R^*$ can detect poor convergence for discrete parameter spaces, like $\widehat{R}$.

@@

Collectively, the results of \S\ref{sec:discrete_small} and \S\ref{sec:discrete_large} show that $R^*$ is able to detect convergence issues in discrete target distributions. The results of \S\ref{sec:discrete_large} hint that $R^*$ may be superior than $\widehat{R}$ in larger discrete spaces, although further empirical investigation is needed.
